---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Sana Nanlawala"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---

::: {.callout-tip icon="false"}
## Github Repo Link

<https://github.com/stat301-2-2024-winter/final-project-2-SanaNanlawala2026.git>
:::

## Data source

This dataframe was sourced from Kaggle at: <https://www.kaggle.com/datasets/dbdmobile/myanimelist-dataset/>

## Why this data

I chose this data set because as someone who interested in anime myself, I would like to see a predictive model built around the ratings that were given to the animes. I think there are a lot of examples of similar models surrounding movies/shows in the West, like with the Netflix Prize dataset for modelling user ratings predictions. It would be interesting to see how it would hold up for anime, which has an entirely demographic of viewers/audience, especially in age.

## Objective/Research Question 
What I will be looking specifically is to create regression models to help predict the anime 'score' that is the outcome variable of my choosing in this dataset. Some predictors will include duration, rating, popularity, favorites, watching, on-hold. 

## Data quality & complexity check

Below, I am reading in the dataset named anime. There are currently 25 variables, of which 9 are numeric and the other character/categorical. I will not be using all 25 variables, as many of them are unrelated or irrelevant such as the id, and the synopsis of the show. However, below I will show in the general reading in of the dataset (excluding the synopsis as it makes the table more readable) 

## Data Import/Reading In
```{r, echo=FALSE}
library(tidyverse)
library(knitr)
anime <- read_csv("data/anime-filtered.csv")
anime|> 
  select(!sypnopsis) |> 
  slice_head(n = 5) |> 
  kable()
```

## Data Quality
To clean up this data further, I will be only evaluating animes that are tv shows specifically, which will then require me to adjust the factors for the duration variable to simply be in minutes (as movies had previously shown hours as well), and will then become a numeric variable. I will also omit some character variables that are irrellevant such as the Japanese and English names of the anime, as the original name is just fine on its own. I will also be excluding the anime_id. 


## Potential data issues

Some potential data issues I can recognize right off the bat is that some shows are listed separately by their season title. When I was looking at the top scored animes, season 3 of Attack on Titan was listed as one of the top, not its entirety of the show. This could be caused my a multitude of reasons, such as because the producing company has changed, for example. Also, another issue shows that most shows are rated at a score of 6.51 which might pose an issue on the quality. 

## Examination of the Target Variable
```{r, echo = FALSE}
library(ggplot2)
library(tidyverse)
anime |> 
  ggplot(aes(Score))+
  geom_bar()
```
A closer look between scores of 6 and 7. 
```{r, echo = FALSE}
anime |> 
  filter(Score > 6 & Score < 7) |> 
  ggplot(aes(Score))+
  geom_bar()
```

Here is a better look at the distribution outside of the score 6.51. The scores are centered around 6.2 and is left skewed.
```{r, echo = FALSE}
anime |> 
  filter(Score != 6.51) |> 
  ggplot(aes(Score))+
  geom_bar()
```

